---
<<<<<<< HEAD
title: "Hawaiian Land Use and Watershed Maps"
author: "Peter Omasta"
=======
title: "Part3"
author: "Me"
>>>>>>> 4d2fd8610485bbd558a6ac8a00b0c9eed954317b
date: "2/19/2020"
output: html_document
---

```{r setup, include=FALSE}
<<<<<<< HEAD
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
#attach necessary packages
library(sf)
library(tmap)
library(tmaptools)
=======
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(sf)
library(tmap)
>>>>>>> 4d2fd8610485bbd558a6ac8a00b0c9eed954317b
library(mapview)
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
<<<<<<< HEAD
library(shinyjs)
=======
>>>>>>> 4d2fd8610485bbd558a6ac8a00b0c9eed954317b
library(paletteer)
```

### **Part 3** 

<<<<<<< HEAD
This will be creating a map which shows he different land use and covers for the Hawaiian islands. It also displays the different watersheds for all of the Hawaiian islands as well. The data was collected by the Hawaii Statewide GIS Program.

```{r, include = FALSE}
#read in the necessary shape files 

land_use <- read_sf(dsn = here::here("data"), layer ="Land_Use_Land_Cover_LULC") %>% 
  clean_names() %>% 
  st_transform(crs = 4326)

#lets just see what we are working with
mapview(land_use)
plot(land_use)

#read in watershed data
water_sheds <- read_sf(dsn = here::here("data"), layer="Watersheds") %>% 
  clean_names()
```

```{r, include= FALSE}
#tidy up the data set that we are working with
land_clean <- land_use %>%  
  mutate(landcover = recode(landcover, 
                            'Orchards, Groves, Vineyards, Nurseries and Ornamental Horticultural Areas' = "Orchards")) %>% 
  filter(landcover %in% c("Orchards",
                          "Residential",
                          "Industrial",
                          "Confined Feeding Operations",
                          "Industrial and Commercial Complexes",
                          "Transitional Areas",
                          "Mixed Urban or Built-up Land",
                          "Strip Mines, Quarries, and Gravel Pits",
                          "Other Agricultural Land",
                          "Transportation, Communications and Utilities",
                          "Reservoirs",
                          "Mixed Rangeland",
                          "Other Urban or Built-up Land",
                          "Commercial and Services",
                          "Cropland and Pasture"
                          ))
#uglyname that needs to be shortened 

ggplot(data = land_use) +
  geom_sf(color = "black",
          size = 0.1) +
  geom_sf(data = land_clean,
          aes(fill= landcover),
          alpha=0.5,
          color="NA",
          show.legend = FALSE) +
  coord_sf(xlim = c(-156, -159), ylim = c(20,22)) 
#hmm this is nice but would be better as an interactive map

```

```{r}
#going to use tmap 
tmap_mode("view") #makes it interactive when knitted 
tm_shape(land_clean) +
  tm_fill("landcover",
          alpha = 0.7,
          style = "fisher",
          palette = "plasma") +
  tm_basemap("Esri.WorldImagery") +
  tm_layout(main.title = "Human Land Use in Hawaii",
            main.title.position = "center",
            main.title.color = "darkblue",
            legend.outside = TRUE)

#much better
```
The map above shows the different land cover types that have been changed to accomodate human use. Land use types include land that has been converted for human habitation (Residential), agriculture (Cropland and Pastureland, Orchards, etc.) and industrial (Industrial, Strip Mines, etc.) 
### **Watersheds**

```{r, include=FALSE}
#time for the watershed maps, start with a basic view

mapview(water_sheds)

#cool, lets just focus on the island of O'ahu and make another interactive tmap

```

```{r}
#map for the watersheds
tm_shape(water_sheds,
         name = "Hawaiian Watersheds")+
    tm_fill("area_sqmi",
          alpha = 0.7,
          style = "fisher",
          palette = "cividis",
          title="Watershed Size Sq. Miles",
          n=7) +
  tm_basemap("Esri.WorldImagery") +
  tm_layout(
    title = "Hawaii Watersheds",
    title.color = "black",
    compass.type = "arrow",
    legend.outside = TRUE,
    legend.outside.position = "left",
    main.title = "Hawaiian Watersheds",
    main.title.color = "black"
  )
```

The above map shows the different watersheds in hawaii, the color corresponds to the size of the watersheds in sq. miles.

### **Citations**

Watersheds. (n.d.). Retrieved February 19, 2020, from http://geoportal.hawaii.gov/datasets/watersheds
=======
This will be creating a map which shows he different land use and covers for the Hawaiian islands. It also displays the different watersheds for all of the Hawaiian islands as well. 

```{r}
#read in the necessary shape files 

land_use <- read_sf(dsn = here::here("data"), layer ="Land_Use_Land_Cover_LULC") %>% 
  clean_names() 

#read in watershed data
water_sheds <- read_sf(dsn = here::here("data"), layer="Watersheds")
```

>>>>>>> 4d2fd8610485bbd558a6ac8a00b0c9eed954317b
